Self Regulation Ontology Retest Data Report

### 2017-03-16 SRO_Report.Rmd
working on new report

---
title: 'Self Regulation Ontology Retest Data: Meaningful Variables Clean'
output:
html_document:
toc: yes
toc_float: yes
---

```{r, message=FALSE, warning=FALSE, include=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(gridExtra)
library(lme4)
library(GGally)
library(jsonlite)
library(lavaan)
library(semTools)
library(psych)
library(GPArotation)
library(rmarkdown)
sem <- function(x) {sd(x, na.rm=T) / sqrt(length(x))}
render_this <- function(){rmarkdown::render('SRO_Meaningful_Vars.Rmd', html_notebook(toc = T, toc_float = T))}
options(digits = 4)
```

# Loading and matching datasets

### Load original data (only cleaned meaningful variables)
```{r}
test_data <- read.csv('/Users/zeynepenkavi/Documents/PoldrackLabLocal/Self_Regulation_Ontology/Data/Complete_01-31-2017/meaningful_variables_clean.csv')
```

### Load retest data (all variables; will subset and clean based on cleaned variables for test data)
```{r}
retest_data <- read.csv('/Users/zeynepenkavi/Documents/PoldrackLabLocal/Self_Regulation_Ontology/Data/Retest_02-11-2017/variables_exhaustive.csv')
```

Function to process lookup table
```{r}
process_lookup <- function(lookup_json){
  lookup_json <- data.frame(unlist(lookup_json))
  lookup_json$sub_id <- as.character(row.names(lookup_json))
  row.names(lookup_json) <- seq(1:nrow(lookup_json))
  names(lookup_json)[1] <- 'worker_id'
  lookup_json$worker_id <- as.character(lookup_json$worker_id)
  lookup_json$complete <- ifelse(lookup_json$worker_id == lookup_json$sub_id, 0, 1)
  lookup_json <- lookup_json %>% arrange(-complete, sub_id)
  return(lookup_json)
}
```

### Extract retest participants from test data

```{r}
test_worker_lookup <- fromJSON('/Users/zeynepenkavi/Documents/PoldrackLabLocal/Self_Regulation_Ontology/Data/Retest_02-11-2017/Local/worker_lookup.json')

#Process full lookup tables for both datasets
test_worker_lookup <- process_lookup(test_worker_lookup)

#Process sub_id columns in the data dataframes
retest_data$X <- as.character(retest_data$X)
test_data$X <- as.character(test_data$X)

@@

#Trim lookup table to only include those we have data for
test_worker_lookup <- test_worker_lookup[test_worker_lookup$sub_id %in% test_data$sub_id,]

#Replace worker_id's in retest_data with original sub_id's 
correct_subids <- function(row, lookup_table){
  sub_id = row$sub_id
  if(sub_id %in% lookup_table$worker_id){
    index = which(sub_id == lookup_table$worker_id)
    sub_id = lookup_table$sub_id[index]
    row$sub_id = sub_id
  }
  return(row)
}

retest_data <- retest_data %>%
  group_by(sub_id) %>%
  do(correct_subids(.,test_worker_lookup))
```

### Checks and matching issues

Are all retest subjects in test_data? No. 
```{r}
sum(retest_data$sub_id %in% test_data$sub_id) == nrow(retest_data)
```

Who is missing?
```{r}
retest_data$sub_id[which(retest_data$sub_id %in% test_data$sub_id == FALSE)]
```

One retest participant is not in test data. Remove that subject from retest_data. This participant should not have completed the retest battery to begin with because they had failed our previous criterion on number of properly completed trials. The first HIT I posted, however, had an issue with the qualifications and before I could terminate it a handful of subjects had accepted and begun the battery so we had to let them finish. So this subject has given us good data for the retest battery but worse data for the test battery.
```{r}
retest_data <- retest_data[retest_data$sub_id %in% test_data$sub_id,]

retest_data
```

Extract test data for retest subjects
```{r}
retest_subs_test_data <- test_data[test_data$sub_id %in% retest_data$sub_id,]

@@

```{r echo=FALSE}
rm(test_worker_lookup, correct_subids, process_lookup)
```

Get only cleaned meaningful var columns based on test data from retest data

```{r}
#Get list of meaningful vars
meaningful_vars = names(test_data)

#Remove transofrmation suffices to be able to use for subsetting of retest data
meaningful_vars = gsub(".logTr|.ReflogTr", "", meaningful_vars)

#Get list of transformed variables and the sign of their skew
logTr_vars = data.frame(grep('logTr', names(test_data), value=T))
names(logTr_vars) = c('var')
logTr_vars$sign = ifelse(grepl(".ReflogTr",logTr_vars$var), 'negative', 'positive')
logTr_vars$var = gsub(".logTr|.ReflogTr", "", logTr_vars$var)

#Subset retest data
retest_data = retest_data[, names(retest_data) %in% meaningful_vars]
```

@@

# Correlations for all cleaned meaningful variables
Spearman only

@@

## Distributions of correlations by task vs. survey

@@

### Rank order of correlations

#### For tasks
```{r}
cor_df %>%
  filter(task == 'task') %>%
  select(dv, spearman) %>%
  arrange(-spearman)
```

#### For surveys

@@

# Different task variable types

@@

## Drift diffusion variables: raw vars vs. EZ vs. hddm

@@

# Replicating the factor structure

## Confirm comparability of retest and rest sample

@@

## Efa and Cfa

>>>>

---
title: 'Self Regulation Ontology Retest Data Report'
output:
html_document:
toc: yes
toc_float: yes
---

```{r, message=FALSE, warning=FALSE, include=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(gridExtra)
library(lme4)
library(GGally)
library(jsonlite)
library(lavaan)
library(semTools)
library(psych)
library(GPArotation)
library(rmarkdown)
sem <- function(x) {sd(x, na.rm=T) / sqrt(length(x))}
render_this <- function(){rmarkdown::render('SRO_Report.Rmd', html_notebook(toc = T, toc_float = T))}
options(digits = 4)
```

# Introduction

# Methods

## Loading and matching datasets

Using the variables that were designated as meaningful before. This is not the smallest subset where certain variables are dropped because they are correlated with other variables and includes both accuracies as well as both kinds of DDM variables (EZ and HDDM). It will not include variables that were considered not of theoretical interest in the analyses of time 1 data. The variables that are not included will be listed for reference as well.

### Load time 1 data
```{r}
test_data <- read.csv('/Users/zeynepenkavi/Documents/PoldrackLabLocal/Self_Regulation_Ontology/Data/Complete_01-31-2017/meaningful_variables.csv')

test_data2 <- read.csv('/Users/zeynepenkavi/Documents/PoldrackLabLocal/Self_Regulation_Ontology/Data/Complete_01-31-2017/meaningful_variables_noDDM.csv')

test_data3 <- read.csv('/Users/zeynepenkavi/Documents/PoldrackLabLocal/Self_Regulation_Ontology/Data/Complete_01-31-2017/meaningful_variables_EZ.csv')

test_data <- merge(test_data, test_data2)
test_data <- merge(test_data, test_data3)
rm(test_data2, test_data3)
```

For reference here are the variables that are not included because they were not of theoretical interest in factor structure analyses of this data so far.

```{r}
test_data2 <- read.csv('/Users/zeynepenkavi/Documents/PoldrackLabLocal/Self_Regulation_Ontology/Data/Complete_01-31-2017/variables_exhaustive.csv')

test_data2$X <- as.character(test_data2$X)

names(test_data2)[1] <- 'sub_id'

df <- data.frame(names(test_data2)[which(names(test_data2) %in% names(test_data) == FALSE)])
names(df) = c('vars')

df
```

```{r echo=FALSE}
rm(test_data2, df)
```

### Load time 2 data 
```{r}
retest_data <- read.csv('/Users/zeynepenkavi/Documents/PoldrackLabLocal/Self_Regulation_Ontology/Data/Retest_02-11-2017/meaningful_variables.csv')

retest_data2 <- read.csv('/Users/zeynepenkavi/Documents/PoldrackLabLocal/Self_Regulation_Ontology/Data/Retest_02-11-2017/meaningful_variables_noDDM.csv')

retest_data3 <- read.csv('/Users/zeynepenkavi/Documents/PoldrackLabLocal/Self_Regulation_Ontology/Data/Retest_02-11-2017/meaningful_variables_EZ.csv')

retest_data <- merge(retest_data, retest_data2)
retest_data <- merge(retest_data, retest_data3)
rm(retest_data2, retest_data3)
```

Extract retest participants from test data

```{r}
#Process sub_id columns in the data dataframes
retest_data$X <- as.character(retest_data$X)
test_data$X <- as.character(test_data$X)

@@

retest_data <- retest_data[retest_data$sub_id %in% test_data$sub_id,]
```

### Clean t1 data

The cleaning functions here are translated from the python pipeline to mimick the same procedure.

Our cleaning procedures includes two steps:  
1. removing outliers for each variables where outliers are defined as those who are greater than 2.5 IQR from median  
2. Log transforming variables with skew. In cleaning T1 data for the `meaningful_variables_clean.csv` variables for which skew could not be removed successfully were dropped. **This step is excluded here. Variables that are not successfully transformed will be noted but not dropped.**

```{r}
remove_outliers = function(data_column, quantile_range = 2.5){

"%w/o%" <- function(x, y) x[!x %in% y]

neg_log <- function(column){
  col_max = max(column, na.rm=T)
  column = col_max+1-column
  return(log(column))
}

transform_remove_skew = function(data, columns, threshold = 1){

  tmp = as.data.frame(apply(data[,columns],2,skew))
  names(tmp) = c("skew")
  tmp$dv = row.names(tmp)
  tmp = tmp %>% 
    filter(abs(skew)>threshold)

  skewed_variables = tmp$dv
  skew_subset = data[, skewed_variables]
  positive_subset = data[,tmp$dv[tmp$skew>0]]
  negative_subset = data[,tmp$dv[tmp$skew<0]]

  # transform variables
  # log transform for positive skew
  positive_subset = log(positive_subset)
  successful_transforms = as.data.frame(apply(positive_subset, 2, skew))
  names(successful_transforms) = c('skew')
  successful_transforms$dv = row.names(successful_transforms)
  successful_transforms = successful_transforms %>% filter(abs(skew)<threshold)
  successful_transforms = successful_transforms$dv
  successful_transforms = positive_subset[,successful_transforms]
  dropped_vars = names(positive_subset) %w/o% names(successful_transforms)

  # replace transformed variables
  data = data[,-c(which(names(data) %in% names(positive_subset)))]
  names(successful_transforms) = paste0(names(successful_transforms), '.logTr')
  cat(rep('*', 40))
  cat('\n')
  cat(paste0('Dropping ', length(dropped_vars) ,' positively skewed data that could not be transformed successfully:'))
  cat('\n')
  cat(dropped_vars, sep = '\n')
  cat(rep('*', 40))
  cat('\n')
  data = cbind(data, successful_transforms)

  # reflected log transform for negative skew      
  negative_subset = as.data.frame(apply(negative_subset, 2, neg_log))
  successful_transforms = as.data.frame(apply(negative_subset, 2, skew))
  names(successful_transforms) = c('skew')
  successful_transforms$dv = row.names(successful_transforms)
  successful_transforms = successful_transforms %>% filter(abs(skew)<1)
  successful_transforms = successful_transforms$dv
  successful_transforms = negative_subset[,successful_transforms]
  dropped_vars = names(negative_subset) %w/o% names(successful_transforms)

  # replace transformed variables
  data = data[,-c(which(names(data) %in% names(negative_subset)))]
  names(successful_transforms) = paste0(names(successful_transforms), '.ReflogTr')
  cat(rep('*', 40))
  cat('\n')
  cat(paste0('Dropping ', length(dropped_vars) ,' negatively skewed data that could not be transformed successfully:'))
  cat('\n')
  cat(dropped_vars, sep = '\n')
  cat(rep('*', 40))
  cat('\n')
  data = cbind(data, successful_transforms)

  return(data)
}
```

------------------------------------------------------------------------
Extract test data for retest subjects
```{r}
retest_subs_test_data <- test_data[test_data$sub_id %in% retest_data$sub_id,]

@@

@@

# Results

## Correlations for all cleaned meaningful variables
Spearman only

@@

### Distributions of correlations by task vs. survey

@@

#### Rank order of correlations

##### For tasks
```{r}
cor_df %>%
  filter(task == 'task') %>%
  select(dv, spearman) %>%
  arrange(-spearman)
```

##### For surveys

@@

## Different task variable types

@@

### Drift diffusion variables: raw vars vs. EZ vs. hddm

**DO THIS ON VARIABLES EXHAUSTIVE GETTING ALL RT AND ACC'S**

@@

## Replicating the factor structure

### Confirm comparability of retest and rest sample

@@

### Efa and Cfa

**DO ONLY ON SURVEY AND TASK DATA SEPARATELY**
